---
title: "Noise"
author: "Daniel K Baissa"
date: "2022-11-03"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(raster)
library(sf)
```

## Does my new idea fight noise?

Let's find out. Let's start by creating a raster where we know the answers. Then we will randomly add some noise.


```{r}
r <- raster(ncol=100, nrow=100)
ncell(r)
hasValues(r)
set.seed(0)
# v <- c(rep(.4, ncell(r)/2),rep(.2, ncell(r)/2))
vals <- rpois(ncell(r), 2)
# v <- abs(rnorm(ncell(r)))
vals<- vals/max(vals)  
values(r)<- vals
# hasValues(r)
# inMemory(r)
# values(r)[1:10]
pdf(file = "Noise_Example/Base_Data.pdf",   # The directory you want to save the file in
    width = 6, # The width of the plot in inches
    height = 4) # The height of the plot in inches
plot(r, main= paste0('Data Simulated from Poison Distribution Lamda = 2'))
dev.off()
```

```{r}
ppside <- 50

h <- ceiling(ncol(r)/ppside)
v <- ceiling(nrow(r)/ppside)

agg <- aggregate(r,fact=c(h,v))
agg[]    <- 1:ncell(agg)
agg_poly <- rasterToPolygons(agg)

plot(agg_poly)
```

Now let's see what the average of each area is

```{r}
base_grid <- agg_poly
r.vals <- raster::extract(r, base_grid)
    # Use list apply to calculate mean for each polygon
    r.mean <- lapply(r.vals, FUN=mean, na.rm = T)
    # Join mean values to polygon data
    base_grid@data <- data.frame(base_grid@data, base=unlist(r.mean))
    
    base_grid = st_as_sf(base_grid)     

```

Now let's add noise to the raster. Let's assume that the noise is Gaussian with mean zero.

```{r}

r2 <- r + rnorm(ncell(r), 0,sd(vals))
# r2 <- r + rnorm(ncell(r), 0,1)

pdf(file = "Noise_Example/Noise_added.pdf",   # The directory you want to save the file in
    width = 6, # The width of the plot in inches
    height = 4) # The height of the plot in inches
plot(r2-r)
dev.off()

```

```{r}
base_raster_diff <- r2-r
base_raster_diff <- as.data.frame(base_raster_diff)

pdf(file = "Noise_Example/Base_Difference.pdf",   # The directory you want to save the file in
    width = 6, # The width of the plot in inches
    height = 4) # The height of the plot in inches

base_raster_diff |> 
  ggplot()+
  geom_density(mapping =  aes(x = layer, y = ..density..)) +
  geom_vline(xintercept = mean(base_raster_diff$layer), colour="red")+
  geom_vline(xintercept = c(sd(base_raster_diff$layer), -1*sd(base_raster_diff$layer)), colour="red", linetype = "longdash")+
    geom_vline(xintercept = c(2*sd(base_raster_diff$layer), -2*sd(base_raster_diff$layer)), colour="green", linetype = "longdash")+
  theme_classic()+
  labs(title = "Difference Between True Raster and True Raster with Noise",
    subtitle = paste("2 Standard Deviations = ", round(2*sd(base_raster_diff$layer), 3)))

dev.off()
```


Now let's do something like a spatial mean filter

```{r}
test_grid <- agg_poly
r.vals <- raster::extract(r2, test_grid)
    # Use list apply to calculate mean for each polygon
    r.mean <- lapply(r.vals, FUN=mean, na.rm = T)
    # Join mean values to polygon data
    test_grid@data <- data.frame(test_grid@data, test=unlist(r.mean))
    
    merged = st_as_sf(test_grid)     

```

```{r}
merged <- cbind(base_grid, test = merged$test)
```

```{r}
merged <- merged |> 
  mutate(diff = base - test) |> 
  mutate(prop_diff = diff/base) |> 
  mutate(prop_diff_test = test/base)
```

```{r}
ggplot()+
  geom_sf(data = merged, mapping =  aes(color = NULL, fill = prop_diff))

ggplot()+
  geom_sf(data = merged, mapping =  aes(color = NULL, fill = prop_diff_test))
```

Now Let's look at the distribution

```{r}
pdf(file = "Noise_Example/Spatial_Mean_Difference.pdf",   # The directory you want to save the file in
    width = 6, # The width of the plot in inches
    height = 4) # The height of the plot in inches


ggplot(data = merged)+
  geom_density(mapping =  aes(x = diff, y = ..density..)) +
  geom_vline(xintercept = mean(merged$diff), colour="red")+
  geom_vline(xintercept = c(sd(merged$diff), -1*sd(merged$diff)), colour="red", linetype = "longdash")+
    geom_vline(xintercept = c(2*sd(merged$diff), -2*sd(merged$diff)), colour="green", linetype = "longdash")+
  theme_classic()+
  labs(title = "Difference Between True Raster and Spatial Mean with Noise", subtitle = paste("2 Standard Deviations = ", round(2*sd(merged$diff), 3)))
dev.off()
```

###

So there is a nice improvement. Now let's stack them on top of this.

```{r}
r3 <- r + rnorm(ncell(r), 0,sd(vals))
r4 <- r + rnorm(ncell(r), 0,sd(vals))
r5 <- r + rnorm(ncell(r), 0,sd(vals))
r6 <- r + rnorm(ncell(r), 0,sd(vals))
r7 <- r + rnorm(ncell(r), 0,sd(vals))

s <- stack(r3,r4,r5,r6,r7)

s[[1,]]
```

```{r}
for (i in 1:5) {
  # r.vals <- raster::extract(paste0("r",i), test_grid)
    r.vals <- raster::extract(s[[i,]], test_grid)

    # Use list apply to calculate mean for each polygon
    r.mean <- lapply(r.vals, FUN=mean, na.rm = T)
    # Join mean values to polygon data
    test_grid@data <- data.frame(test_grid@data, test=unlist(r.mean))
    colnames(test_grid@data)[length(colnames(test_grid@data))] <- paste0("test", i)
    
}
  test_grid = st_as_sf(test_grid)   
```


```{r}
test_grid <- cbind(test_grid, base = merged$base)

test_grid <- test_grid |> 
  group_by(layer) |> 
  mutate(mean_test = mean(c(test,test1,test2,test3,test4,test5))) |> 
  mutate(stack_diff = (mean_test - base))
         

```


```{r}
ggplot()+
  geom_sf(data = test_grid, mapping =  aes(color = NULL, fill = mean_test))

ggplot()+
  geom_sf(data = test_grid, mapping =  aes(color = NULL, fill = stack_diff))
```

```{r}
pdf(file = "Noise_Example/Stacked_Spatial_Mean_Difference.pdf",   # The directory you want to save the file in
    width = 6, # The width of the plot in inches
    height = 4) # The height of the plot in inches


ggplot(data = test_grid)+
  geom_density(mapping =  aes(x = stack_diff, y = ..density..)) +
  geom_vline(xintercept = mean(test_grid$stack_diff), colour="red")+
  geom_vline(xintercept = c(sd(test_grid$stack_diff), -1*sd(test_grid$stack_diff)), colour="red", linetype = "longdash")+
    geom_vline(xintercept = c(2*sd(test_grid$stack_diff), -2*sd(test_grid$stack_diff)), colour="green", linetype = "longdash")+
  theme_classic()+
  labs(title = "Difference Between True Raster and 6 Stacked Rasters \nwith Spatial Mean and Noise", subtitle = paste("2 Standard Deviations = ", round(2*sd(test_grid$stack_diff), 3)))

dev.off()
```



```{r}
pdf(file = "Noise_Example/Comparison_Difference.pdf",   # The directory you want to save the file in
    width = 6, # The width of the plot in inches
    height = 4) # The height of the plot in inches

ggplot()+
  geom_density(data = test_grid, mapping =  aes(x = stack_diff, y = ..density..) , fill = "red", alpha =.5) +
  geom_density(data = merged, mapping =  aes(x = diff, y = ..density..),fill = "blue", alpha =.5) +
  geom_density(data = base_raster_diff, mapping =  aes(x = layer, y = ..density..),fill = "green", alpha =.5) +
  theme_classic()+
  xlab("Error") + ylab("Density")
dev.off()
```

```{r}
t <- rbinom(10, 1,.25)
t[which(t >0 )] <- rnorm(length(t[which(t >0 )]), 0,sd(vals))

t
```

```{r}
t <- rbinom(ncell(r), 1,.25)
t[which(t >0 )] <- rnorm(length(t[which(t >0 )]), 0,sd(vals))

rt1 <- r + t

```

```{r}
base_raster_diff <-r- rt1
base_raster_diff <- as.data.frame(base_raster_diff)

base_raster_diff |> 
  ggplot()+
  geom_density(mapping =  aes(x = layer, y = ..density..)) +
  geom_vline(xintercept = mean(base_raster_diff$layer), colour="red")+
  geom_vline(xintercept = c(sd(base_raster_diff$layer), -1*sd(base_raster_diff$layer)), colour="red", linetype = "longdash")+
    geom_vline(xintercept = c(2*sd(base_raster_diff$layer), -2*sd(base_raster_diff$layer)), colour="green", linetype = "longdash")+
  theme_classic()+
  labs(title = paste("2 Standard Deviations = ", round(2*sd(base_raster_diff$layer), 3)))
```
```{r}
ppside <- 50

h <- ceiling(ncol(r)/ppside)
v <- ceiling(nrow(r)/ppside)

agg <- aggregate(r,fact=c(h,v))
agg[]    <- 1:ncell(agg)
agg_poly <- rasterToPolygons(agg)

plot(agg_poly)
```

```{r}
test_grid <- agg_poly
r.vals <- raster::extract(rt1, test_grid)
    # Use list apply to calculate mean for each polygon
    r.mean <- lapply(r.vals, FUN=mean, na.rm = T)
    # Join mean values to polygon data
    test_grid@data <- data.frame(test_grid@data, test.rand=unlist(r.mean))
    
    test_grid = st_as_sf(test_grid)  

merged <- cbind(merged, test_grid)
merged <- merged |> 
  mutate(diff = base - test.rand) 
    
ggplot(data = merged)+
  geom_density(mapping =  aes(x = diff, y = ..density..)) +
  geom_vline(xintercept = mean(merged$diff), colour="red")+
  geom_vline(xintercept = c(sd(merged$diff), -1*sd(merged$diff)), colour="red", linetype = "longdash")+
    geom_vline(xintercept = c(2*sd(merged$diff), -2*sd(merged$diff)), colour="green", linetype = "longdash")+
  theme_classic()+
  labs(title = paste("2 Standard Deviations = ", round(2*sd(merged$diff), 3)))
```

```{r}

t <- rbinom(ncell(r), 1,.25)
t[which(t >0 )] <- rnorm(length(t[which(t >0 )]), 0,sd(vals))

rt3 <- r + t

t <- rbinom(ncell(r), 1,.25)
t[which(t >0 )] <- rnorm(length(t[which(t >0 )]), 0,sd(vals))

rt4 <- r + t
t <- rbinom(ncell(r), 1,.25)
t[which(t >0 )] <- rnorm(length(t[which(t >0 )]), 0,sd(vals))

rt5 <- r + t
t <- rbinom(ncell(r), 1,.25)
t[which(t >0 )] <- rnorm(length(t[which(t >0 )]), 0,sd(vals))

rt6 <- r + t
s <- stack(rt1, rt3,rt4,rt5,rt6)

s[[1,]]
```

```{r}
test_grid <- agg_poly
for (i in 1:5) {
  # r.vals <- raster::extract(paste0("r",i), test_grid)
    r.vals <- raster::extract(s[[i,]], test_grid)

    # Use list apply to calculate mean for each polygon
    r.mean <- lapply(r.vals, FUN=mean, na.rm = T)
    # Join mean values to polygon data
    test_grid@data <- data.frame(test_grid@data, test=unlist(r.mean))
    colnames(test_grid@data)[length(colnames(test_grid@data))] <- paste0("test.rand", i)
    
}
  test_grid = st_as_sf(test_grid)  
  
  test_grid <- cbind(test_grid, merged$base)

```

```{r}
test_grid <- test_grid |> 
  group_by(layer) |> 
  mutate(mean_test = mean(c(test.rand1,test.rand2,test.rand3,test.rand4,test.rand5))) |> 
  mutate(stack_diff = (mean_test - merged.base))

ggplot(data = test_grid)+
  geom_density(mapping =  aes(x = stack_diff, y = ..density..)) +
  geom_vline(xintercept = mean(test_grid$stack_diff), colour="red")+
  geom_vline(xintercept = c(sd(test_grid$stack_diff), -1*sd(test_grid$stack_diff)), colour="red", linetype = "longdash")+
    geom_vline(xintercept = c(2*sd(test_grid$stack_diff), -2*sd(test_grid$stack_diff)), colour="green", linetype = "longdash")+
  theme_classic()+
  labs(title = paste("2 Standard Deviations = ", round(2*sd(test_grid$stack_diff), 3)))
```

